### 2014-06-16 Equipment/Equipment_Guide/Equipment_Guide.tex
Equipment Guide: Edits to water flow sensor notes

>>>>

\chapter{Water Flow Sensor}
\label{chap:Water_Flow_Sensor}

\section{Arduino Config, Setup, and Data Logging}

The Arduinos record voltage data locally, and they transmit voltage data to the host computer via the AMQP messaging protocol. The host computer uses RabbitMQ to act as a broker. Python scripts and the pika module send and receive AMQP messages. The following steps setup the host computer and Arduinos for this purpose.

\subsection{Host Computer (RabbitMQ)}

Note: the following assumes that the server/host computer's address is 192.168.1.100. It's best to configure the host computer with a static IP address using static DHCP or manually.

\begin{enumerate}
\item Install RabbitMQ for your platform and start the RabbitMQ server (./sbin/rabbitmq-server on Mac or Linux).
https://www.rabbitmq.com/
\item Copy the receive\_data.py script from the NIST-FIRE repository to the data logging computer.
\item Install Python (The Anaconda Python distribution from Continuum Analytics is recommended)
\item Install the pika module for Python using
\begin{verbatim}
pip install pika
\end{verbatim}
\item Run the receive\_data.py script as
\begin{verbatim}
python receive_data.py 192.168.1.100 output.csv
\end{verbatim}
\end{enumerate}

\subsection{Host Computer (NTP Time Server)}

On Windows, download and install the NTP server

http://www.meinbergglobal.com/english/sw/ntp.htm

Note: you might need to disable the Windows Firewall service so that incoming connections will not be blocked.

Mac and Linux use NTP natively.

\subsection{Arduino Client (Sketch)}

\begin{enumerate}
\item Install Arduino IDE
http://arduino.cc/
\item Plug the Arduino Yun into the USB port of the host computer.
\item Set the Tools > Board to Arduino Yun and the Tools > Port to the /dev/ttyusbmodemXXXXXX setting that matches the plugged-in Arduino.
\item A Sketch is a script that runs on the Arduino microcontroller. Using the Arduino IDE, install the pressure\_sensor Sketch on the Arduino from the NIST-FIRE repository. This enables the reading and writing of data from the analog-to-digital converter via the REST API on the Arduino. Note: this sketch requires the Adafruit ADS 1015 library from https://github.com/adafruit/Adafruit\_ADS1X15.
\end{enumerate}

\subsection{Arduino Client (Linux)}

\begin{enumerate}
\item Unplug the Arduino from the computer, and power on the Arduino from the battery.
\item Connect to the Arduino WiFi network. Access the Arduino web interface (http://arduino.local) with the default password of ``arduino''. Set the Arduino name (referred to as <arduino\_name> in the following steps), time zone, and WiFi network information on the Arduino web config. Also, set the ``REST API Access'' to be ``Open''.
\item SSH to the Arduino at root@<arduino\_name>.local and install the SFTP server, ntpclient, pip, and pika (for AMQP) on the Arduino
\begin{verbatim}
opkg update
opkg install openssh-sftp-server ntpclient distribute python-openssl
easy_install pip
pip install pika
\end{verbatim}
\item By default, the microSD card on the Arduino is mounted at /mnt/sda1. Copy the send\_data.py script from the NIST-FIRE repository to /mnt/sda1 on the Arduino.
\item Go to System > Startup > Local Startup in the Arduino web config and add a startup line to the Local Config section (/etc/rc.local)
\small
\begin{verbatim}
python /mnt/sda1/send_data.py 192.168.1.100 <arduino_name> /mnt/sda1/output.csv
\end{verbatim}
\normalsize
\clearpage
\item The default NTP service is not very robust, so we will replace it with ntpclient and have it sync with the NTP server every 2 minutes. Disable the default NTP service using
\begin{verbatim}
/etc/init.d/sysntpd disable
\end{verbatim}
Configure the ntpclient service by editing /etc/config/ntpclient to be
\begin{verbatim}
config ntpserver
        option hostname '192.168.1.100'
        option port     '123'

config ntpserver
        option hostname '0.pool.ntp.org'
        option port     '123'

config ntpserver
        option hostname '1.pool.ntp.org'
        option port     '123'

config ntpserver
        option hostname '2.pool.ntp.org'
        option port     '123'

config ntpserver
        option hostname '3.pool.ntp.org'
        option port     '123'

config ntpdrift                                 
        option freq     '0'

config ntpclient                                
        option interval 120
\end{verbatim}
\item Reset power to the Arduino, and then it should start broadcasting data values within about 2 minutes of booting.
\end{enumerate}

\section{Parts}

\begin{itemize}
\item Mouser \#:   517-929834-02-36-RK \\
Mfr. \#: 929834-02-36-RK \\
Desc.: Headers \& Wire Housings FULL STICK HDR/36POS \\
\$1.13

\item Mouser \#:   932-MIKROE-197 \\
Mfr. \#: MIKROE-197 \\
Desc.:  Daughter Cards \& OEM Boards SMARTPROTO PROTO ADAPTER BOARD \\
\$4.90
\end{itemize}

%  ==============
%  = Appendices =
%  ==============
